ZCB Pricing

### 2018-07-01 ZCB Pricing .Rmd
More on the explicit method

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

alpha = 0.01
beta = 0.1
sigma = 0.04
I = 1:20
J = seq(0,0.2,0.005)

dt = 1/length(I)
dr = J[2] -J[1]

J[8]
nr <- length(J)
nt <- length(I)+1
V <- matrix(NA, nrow = length(J), ncol = length(I)+1)

V[,length(I)+1] <- rep(1)

count <- 0
for (i in rev(I)){
  count <- count+1
for (j in (1+count):(nr-count)){
  pu <- sigma^2/(2*dr^2) + (alpha -beta*J[nr-j])/(2*dr)
  pm <- 1/dt -sigma^2/(dr^2)
  pd <- sigma^2/(2*(dr^2)) - (alpha -beta*J[nr-j])/(2*dr) 
  V[j,i] <- (1/(J[nr-j]+1/dt))*(pu*V[j-1,i+1] + pm*V[j,i+1] +pd*V[j+1,i+1])
}

@@

V[(nr+1)/2,1]

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls(all =TRUE))
```

@@

explicit <- function(alpha,beta,sigma,r,I,dr){

# I = partition of time
# dr = length of rate sub-intervals
# r must be in the middle of partition of J
# length of J must be twice the length of I

J <- seq(0,2*r,dr)

dt = 1/length(I) # length of each time sub-interval
dr = J[2] -J[1] # length of each rate sub-interval

V <- matrix(NA, nrow = length(J), ncol = length(I)+1) # empty matrix
nr <- length(J) # number of rows of the matrix - represents partition of rate
nt <- length(I)+1 # number of columns - represents partition of time

V[,length(I)+1] <- rep(1) # Boundary condition at expiry of bond

count <- 0 # set up count variable
for (i in rev(I)){ # rev is the reverse order of the time partition
  count <- count+1 # keeping track of loop
  for (j in (1+count):(nr-count)){
    pu <- sigma^2/(2*dr^2) + (alpha -beta*J[nr-j])/(2*dr) # up value
    pm <- 1/dt -sigma^2/(dr^2) # middle value
    pd <- sigma^2/(2*(dr^2)) - (alpha -beta*J[nr-j])/(2*dr) # down value
    V[j,i] <- (1/(J[nr-j]+1/dt))*(pu*V[j-1,i+1] + pm*V[j,i+1] +pd*V[j+1,i+1])
  }

@@

V[(nr+1)/2,1]
}

explicit(0.01,0.1,0.04,0.10,1:20,0.005)
